---
title: "pset1"
format: html
---

# KAT Statistics Problem Set #1

## Introduction

Let's use what we know about graphing categorical and quantitative data to explore some real marine biology data compiled from several sources into a single dataset here: https://www.kaggle.com/datasets/maulikgajera/aquatic-wildlife-atlas-global-species-records.

```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
aquatic_animals = read_csv("aquatic_animals_200k.csv") |> clean_names()
```

First, look at the data.

```{r}
aquatic_animals
```

There are lots of variables in this dataset. We can look at a summary of each variable like this:

```{r}
glimpse(aquatic_animals)
```

**Question 1: What do you notice about the variables labeled `<dbl>` versus the variables labeled `<chr>`?**

\[your answer here\]

## Categorical Data

Choose one categorical variable and make a barplot.

(Because we haven't already calculated the frequencies into a separate column, we can use `geom_bar` instead of `geom_col`)

```{r}
aquatic_animals |> 
  ggplot() +
  aes(x = YOUR_CATEGORICAL_VARIABLE_HERE) +
  geom_bar()
```

Now pick two categorical variables and make a stacked barplot.

```{r}
aquatic_animals |> 
  ggplot() +
  aes(x = ONE_CATEGORICAL_VARIABLE, fill = ANOTHER_CATEGORICAL_VARIABLE) +
  geom_bar()
```

**Question 2: Are the two variables that you chose independent or associated? Explain.**

\[your answer here\]

## Quantitative Data

Choose one quantitative variable and make a histogram.

```{r}
aquatic_animals |> 
  ggplot() +
  aes(x = YOUR_QUANTITATIVE_VARIABLE) +
  geom_histogram()
```

**Question 3: Describe the distribution of your quantitative variable:**

Remember to comment on shape, center, spread, outliers, and context.

\[your answer here\]

## Explore

If you have any extra time, feel free to explore different variables in the dataset. If you want to make a new code block for yourself, the keyboard shortcut is Cmd + Shift + i.
